---
title: "Let's R"
---

#### Day 3 of data meetings 

Over the last two meetings we have:  

*  Learned about the metabolomics workflow and learned about how a feature table is created.  
*  We installed tidyverse and had some luck loading other packages (Will move to R studio cloud to analyze metabolomics data)
*  We spoke about project organization and the importance of following a consistent strategy to store all the files the are needed and generated during data analysis  
*  We read about the importance of version control in [this paper by Jennifer Bryan](https://amstat.tandfonline.com/doi/full/10.1080/00031305.2017.1399928?journalCode=utas20#.XnExyJNKi_X)
*  We installed git on our personal computers  
*  We learned about the layout of Rstudio and the IDE environment  
*  We learned about different types of onjects and different classes of vectors  
*  We used simulated data to track worm growth over time  
*  We learned about ggplot and the way graphs are layered when using the ggplot2 package  
*  We used the tidyverse universe and the *piping* ( %>% ) that is characteristic of this universe (chapter 14 in [R4DS book](https://r4ds.had.co.nz/))  
*  We learned about iteration using a for loop (for more information, read chapter 17 in the [R4DS book](https://r4ds.had.co.nz/))  
*  We created an R project to hold our code and reports generated about the simulated worm growth data  
*  We linked Rstudio to git and saw the git tab pop up in the environment pane  

    
Today, we will learn more about R.  

*  We will spend an hour or so talking about R and using R studio.   
    *  Create a function in R  
    *  Continue working with the R project  
    *  Use git to link R studio to github completely and push the project as a repository to github  
    *  Use the daf-16 metabolomics dataset as an example to go over metabolomics data analysis
        *  Import metabolomics data and clean the dataset  
        *  Create a PCA plot  
    
    
*****************

#### Some notes on R  

##### Functions 

**When should you write a function?**  
Consider writing a function whenever you've copied and pasted a block of code more than twice (i.e, you now have three copies of the same code)  

**Three big advantages:**  
1.  You can give a function an evocative name - makes code easier to understand  
2.  If requirements change, only need to update code in one place, instead of many  
3.  Eliminate the chance of making incidental mistakes while copying and pasting  

**Three key steps to making a function:**  
1.  Pick a name  
2.  List the inputs, i.e., the arguments  
3.  Place the code in the body of the funnction  

Syntax for a function:  

```{r func, eval = F}
function_name <- function(arg1 = default1, arg2 = default2, ...){  
                            "body of function"
                    }
```


##### Iteration: For loops and purrr

**Every for loop has the following three components:**  
1.  *output* vector or object with sufficient space allocated for  the output  
2.  *sequence* $i$ determines what to loop over. Each runnn of the for loop will assign $i$ to a different value.  
3.  *body* is the code that does the work  

Syntax for a for loop:  

```{r for loop, eval = F}
output <- vector("double",ncol(df)) #1. output
for(i in seq_along(df)){            #2. sequence
    output[i] <- median(df[i])      #3. body
}
```


In the tidyverse universe, the purrr package is used for iteration over columns in a tibble / dataframe. See chapter 17 in [R4DS book](https://r4ds.had.co.nz/).  


##### Tidyverse terminology  

**Tibble**: Tibbles are dataframes but are special. They differ in the following ways:  

*  Tibbles never change the type of the inputs (e.g.: never converts strinngs to factors)  
*  Tibbles never change the names of variables  
*  Tibbles never create row names  
*  Tibbles can have column names that are not valid R variable names. For example: might not start with a letter, or might contain unnsual characters, like a space. To refer to these variables, will need to surround them with backticks, `.  

See chapter 7 in [R4DS book](https://r4ds.had.co.nz/) for more information  


**************



                    



